Supporting Information:\\ Headgroup structure and cation binding in phosphatidylserine lipid bilayers

### 2019-04-26 Manuscript/manuscriptPSsuppl.tex
Simulation description and author contribution for J. Melcr

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium} \\
%The data in Table \ref{mixedIONsystems} and Refs \citenum{lipid17_cacl_series,POPCpopsLIPID17withCaCl}

One set of simulations of POPC:POPS (5:1) mixtures with additional calcium ions has been performed using Amber18 simulation package \cite{amber18md}. Number of ions and water molecules for this set of simulations are shown in Table~\ref{mixedIONsystems}. The simulation details, minimization, heating, and equilibration runs are the same as described above for the other POPC:POPS (5:1) simulations with Amber Lipid17 forcefield. For all simulations, ff99SB~\cite{aqvist90} ion parameters and Amber lipid 17 force field were used. For each ion concentration, 100~ns equilibration runs were followed by 300~ns production runs with two independent trajectories, of which last 200~ns were used for the analysis. The trajectory files can be found in Ref~\citenum{POPCpopsLIPID17withCaCl}

\todo{J. Melcr, please write simulation details about Gromacs simulations with Dang ions.} \\

@@

\noindent
{\it Josef Melcr}

>>>>

\noindent
{\it POPC:POPS (5:1) mixtures with additional calcium chloride concentration} \\
%The data in Table \ref{mixedIONsystems} and Refs \citenum{lipid17_cacl_series,POPCpopsLIPID17withCaCl} 
We have performed two sets of simulations of POPC:POPS (5:1) mixtures using Amber Lipid17 force field \cite{gould18}
with additional calcium chloride concentrations, 
which employed different simulation engines and parameters for ions.

One set of simulations has been performed using Amber18 simulation package \cite{amber18md} and ff99SB~\cite{aqvist90} ion parameters. 
Number of ions and water molecules for this set of simulations are shown in Table~\ref{mixedIONsystems}. The simulation details, minimization, heating, and equilibration runs are the same as described above for the other POPC:POPS (5:1) simulations with Amber Lipid17 forcefield. 
For each ion concentration, 100~ns equilibration runs were followed by 300~ns production runs with two independent trajectories, of which last 200~ns were used for the analysis. The trajectory files can be found in Ref~\citenum{POPCpopsLIPID17withCaCl}

Another set of simulations with Lipid17 force field for lipids \cite{gould18}
and Dang parameters for ions \cite{smith94, dang06}
was performed using the GROMACS 2018 simulation package~\cite{abraham2015gromacs}. 
We used the time step of 2~fs. 
The non-bonded interactions were calculated directly within 1.0~nm cutoff using the Verlet scheme \cite{Pall13}, 
and the long-range electrostatic forces were calculated using particle mesh Ewald~\cite{essman95}. 
The bond lengths of hydrogen atoms were constrained using LINCS algorithm \cite{hess97}.
Temperature was coupled to the velocity rescaling thermostat \cite{bussi07} at 298~K with a coupling constant of 1~ps.
Pressure was coupled to the Parrinello-Rahman barostat \cite{parrinello81} at 1~bar with a coupling constant of 10~ps. 
The composition of the individual simulations is shown in Table~\ref{mixedIONsystems}. 
Initial structures obtained from CHARMM-GUI~\cite{lee16,jo18} 
were equilibrated prior to production simulations used for analysis,
which are available in Ref.~\citenum{lipid17_cacl_series} including the simulation settings. 
Topologies in a format for the GROMACS package are available in the repository \citenum{ECCpops}.

@@

\noindent
{\it Josef Melcr}
consulted the project, 
prepared tools for calculating order parameters
and performed several simulations.
